Sampling Strategies in Decisions from Experience

### 2021-10-01 manuscript.Rmd
Restructuring of the manuscript + Analysis of Probability weighting function

---
title: 'Sampling Strategies in Decisions from Experience'
author: "Linus Hof, Thorsten Pachur, Veronika Zilker"
bibliography: references_sampling-in-dfe.bib
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
    number_sections: no
  pdf_document:
    toc: yes
csl: apa.csl
editor_options: 
  markdown: 
    wrap: sentence
---

```{r}
# load packages
pacman::p_load(tidyverse,
               knitr, 
               viridis)
```

@@

...

@@

...

Given that

$$\begin{equation}
P\left(\lim_{N_X \to \infty} \overline{X}_{N_X} = E(X) \right) = 
P\left(\lim_{N_Y \to \infty} \overline{Y}_{N_Y} = E(Y) \right) = 
1 \; ,
\end{equation}$$

we obtain that

$$\begin{equation}
\left(
    \frac{\overline{X}_{N_X}} {\overline{Y}_{N_Y}}
\right)
\in S : 
P\left(\lim_{N_X \to \infty} \lim_{N_Y \to \infty} \frac{\overline{X}_{N_X}} {\overline{Y}_{N_Y}} = 
\frac{E(X)} {E(Y)}
\right ) =
1 \; ,
\end{equation}$$

I.e., for the summary strategy, we assume that for increasing sample sizes $N_X$ and $N_Y$, the prospect with the larger expected value is chosen almost surely.

@@

@@

Under each condition, i.e., strategy-parameter combinations, all gambles are played by 100 synthetic agents.
We test a set of gambles, in which one of the prospects contains a safe outcome and the other two risky outcomes (*safe-risky gambles*).
Therefore, 60 gambles from an initial set of 10,000 are sampled.
Both outcomes and probabilities are drawn from uniform distributions, ranging from 0 to 20 for outcomes and from .01 to .99 for probabilities of the lower risky outcomes $p_{low}$.
The probabilities of the higher risky outcomes are $1-p_{low}$, respectively.
To omit dominant prospects, safe outcomes fall between both risky outcomes.
Table A1 contains the test set of 60 gambles [Appendix 2][Appendix 2: Prospect Test Set].
Sampling of gambles was stratified, randomly drawing an equal number of 20 gambles with no, an attractive, and an unattractive rare outcome.
Risky outcomes are considered *"rare"* if their probability is $p < .2$ and *"attractive"* (*"unattractive"*) if they are higher (lower) than the safe outcome.

## Model Parameters

**Switching probability** $s$ is the probability with which agents draw the following single sample from the prospect they did not get their most recent single sample from.
$s$ is varied between .1 to 1 in increments of .1.

The **boundary type** is either the minimum value any prospect's sample statistic must reach (absolute) or the minimum value for the difference of these statistics (relative).
Sample statistics are sums over outcomes (comprehensive strategy) and sums over wins (piecewise strategy), respectively.

For comprehensive integration, the **boundary value** $a$ is varied between 15 to 75 in increments of 15.
For piecewise integration $a$ is varied between 1 to 5 in increments of 1.

```{r message=FALSE}
# read choice data 
cols <- list(.default = col_double(),
             strategy = col_factor(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices <- read_csv("data/choices/choices.csv", col_types = cols)
```

In sum, 2 (strategies) x 60 (gambles) x 100 (agents) x 100 (parameter combinations) = `r nrow(choices)` choices are simulated.

@@

Because we are not interested in deviations from normative choice due to sampling artifacts (e.g., ceiling effects produced by low boundaries), we remove trials in which only one prospect was attended.
In addition, we use relative frequencies of sampled outcomes rather than 'a priori' probabilities to compare actual against normative choice behavior.

```{r}
# remove choices where prospects were not attended
choices <- choices %>%
  filter(!(is.na(a_ev_exp) | is.na(b_ev_exp)))
```

```{r eval = FALSE}
# remove choices where not all outcomes were sampled
choices <- choices %>% 
  filter(!(is.na(a_ev_exp) | is.na(b_ev_exp) | a_p1_exp == 0 | a_p2_exp == 0))
```

Removing the respective trials, we are left with `r nrow(choices)` choices.

## Sample Size

```{r message=FALSE}
samples <- choices %>% 
  group_by(strategy, s, boundary, a) %>% 
  summarise(n_med = median(n_sample))
samples_piecewise <- samples %>% filter(strategy == "piecewise")
samples_comprehensive <- samples %>% filter(strategy == "comprehensive")
```

The median sample sizes generated by different parameter combinations ranged from `r min(samples_piecewise$n_med)` to `r max(samples_piecewise$n_med)` for piecewise integration and `r min(samples_comprehensive$n_med)` to `r max(samples_comprehensive$n_med)` for comprehensive integration.

### Boundary type and boundary value (a)

As evidence is accumulated sequentially, relative boundaries and large boundary values naturally lead to larger sample sizes, irrespective of the integration strategy.

```{r message=FALSE}
group_med <- samples_piecewise %>%
  group_by(boundary, a) %>% 
  summarise(group_med = median(n_med)) # to get the median across all s values

samples_piecewise %>%
  ggplot(aes(a, n_med, color = a)) + 
  geom_jitter(alpha = .5, size = 2) +
  geom_point(data = group_med, aes(y = group_med), size = 3) +
  facet_wrap(~boundary) + 
  scale_color_viridis() + 
  labs(title = "Piecewise Integration",
       x ="a", 
       y="Sample Size", 
       col="a") + 
  theme_minimal()
```

```{r message=FALSE}
group_med <- samples_comprehensive %>%
  group_by(boundary, a) %>% 
  summarise(group_med = median(n_med))

samples_comprehensive %>%
  ggplot(aes(a, n_med, color = a)) + 
  geom_jitter(alpha = .5, size = 2) +
  geom_point(data = group_med, aes(y = group_med), size = 3) +
  facet_wrap(~boundary) + 
  scale_color_viridis() + 
  labs(title = "Comprehensive Integration",
       x ="a", 
       y="Sample Size", 
       col="a") + 
  theme_minimal()
```

### Switching probability (s)

For piecewise integration, there is an inverse relationship between switching probability and sample size.
I.e., the lower s, the less frequent prospects are compared and thus, boundaries are only approached with larger sample sizes.
This effect is particularly pronounced for low probabilities such that the increase in sample size accelerates as switching probability decreases.

```{r message=FALSE}
group_med <- samples_piecewise %>%
  group_by(boundary, s) %>% 
  summarise(group_med = median(n_med)) # to get the median across all a values

samples_piecewise %>%
  ggplot(aes(s, n_med, color = s)) + 
  geom_jitter(alpha = .5, size = 2) +
  geom_point(data = group_med, aes(y = group_med), size = 3) +
  facet_wrap(~boundary) + 
  scale_color_viridis() + 
  labs(title = "Piecewise Integration",
       x ="s", 
       y="Sample Size", 
       col="s") + 
  theme_minimal()
```

For comprehensive integration, boundary types differ in the effects of switching probability.
For absolute boundaries, switching probability has no apparent effect on sample size as the distance of a given prospect to its absolute boundary is not changed by switching to (and sampling from) the other prospect.
For relative boundaries, however, samples sizes increase with switching probability.

```{r message=FALSE}
group_med <- samples_comprehensive %>%
  group_by(boundary, s) %>% 
  summarise(group_med = median(n_med)) # to get the median across all a values

samples_comprehensive %>%
  ggplot(aes(s, n_med, color = s)) + 
  geom_jitter(alpha = .5, size = 2) +
  geom_point(data = group_med, aes(y = group_med), size = 3) +
  facet_wrap(~boundary) + 
  scale_color_viridis() + 
  labs(title = "Comprehensive Integration",
       x ="s",
       y = "Sample Size", 
       col="s") + 
  theme_minimal()
```

## Choice Behavior

Below, in extension to Hills and Hertwig [-@hillsInformationSearchDecisions2010], the interplay of integration strategies, gamble features, and model parameters in their effects on choice behavior in general and their contribution to underweighting of rare events in particular is investigated.
We apply two definitions of underweighting of rare events: Considering false response rates, we define underweighting such that the rarity of an attractive (unattractive) outcome leads to choose the safe (risky) prospect although the risky (safe) prospect has a higher expected value.

```{r message=FALSE}
fr_rates <- choices %>% 
  mutate(ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2), 
         norm = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) %>% 
  filter(!is.na(norm)) %>% # exclude trials with normative indifferent options
  group_by(strategy, s, boundary, a, rare, norm, choice) %>% # group correct and incorrect responses
  summarise(n = n()) %>% # absolute numbers 
  mutate(rate = round(n/sum(n), 2), # response rates 
         type = case_when(norm == "A" & choice == "B" ~ "false safe", norm == "B" & choice == "A" ~ "false risky")) %>% 
  filter(!is.na(type)) # remove correct responses
```

Considering the parameters of Prelec's [-@prelecProbabilityWeightingFunction1998] implementation of the weighting function [CPT; cf. @tverskyAdvancesProspectTheory1992], underweighting is reflected by decisions weights estimated to be smaller than the corresponding objective probabilities.

### False Response Rates

```{r message=FALSE}
fr_rates_piecewise <- fr_rates %>% filter(strategy == "piecewise")
fr_rates_comprehensive <- fr_rates %>% filter(strategy == "comprehensive")
```

The false response rates generated by different parameter combinations ranged from `r min(fr_rates_piecewise$rate)` to `r max(fr_rates_piecewise$rate)` for piecewise integration and from `r min(fr_rates_comprehensive$rate)` to `r max(fr_rates_comprehensive$rate)` for comprehensive integration.
However, false response rates vary considerably as a function of rare events, indicating that their presence and attractiveness are large determinants of false response rates.

```{r message=FALSE}
fr_rates %>% 
  group_by(strategy, boundary, rare) %>% 
  summarise(min = min(rate),
            max = max(rate)) %>% 
  kable()
```

The heatmaps below show the false response rates for all strategy-parameter combinations.
Consistent with our - somewhat rough - definition of underweighting, the rate of false risky responses is generally higher, if the unattractive outcome of the risky prospect is rare (top panel).
Conversely, if the attractive outcome of the risky prospect is rare, the rate of false safe responses is generally higher (bottom panel).
As indicated by the larger range of false response rates, the effects of rare events are considerably larger for piecewise integration.

```{r message=FALSE}
fr_rates %>% 
  filter(strategy == "piecewise", boundary == "absolute") %>% 
  ggplot(aes(a, s, fill = rate)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_tile(colour="white", size=0.25) + 
  scale_x_continuous(expand=c(0,0), breaks = seq(1, 5, 1)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(.1, 1, .1)) +
  scale_fill_viridis() + 
  labs(title = "Piecewise Integration | Absolute Boundary",
       x = "a", 
       y= "s", 
       fill = "% False Responses") + 
  theme_minimal() 
```

```{r message=FALSE}
fr_rates %>% 
  filter(strategy == "piecewise", boundary == "relative") %>% 
  ggplot(aes(a, s, fill = rate)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_tile(colour="white", size=0.25) + 
  scale_x_continuous(expand=c(0,0), breaks = seq(1, 5, 1)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(.1, 1, .1)) +
  scale_fill_viridis() + 
  labs(title = "Piecewise Integration | Relative Boundary",
       x = "a", 
       y= "s", 
       fill = "% False Responses") + 
  theme_minimal() 
```

```{r message=FALSE}
fr_rates %>% 
  filter(strategy == "comprehensive", boundary == "absolute") %>% 
  ggplot(aes(a, s, fill = rate)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_tile(colour="white", size=0.25) + 
  scale_x_continuous(expand=c(0,0), breaks = seq(15, 75, 15)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(.1, 1, .1)) +
  scale_fill_viridis() + 
  labs(title = "Comprehensive Integration | Absolute Boundary",
       x = "a", 
       y= "s", 
       fill = "% False Responses") + 
  theme_minimal() 
```

```{r message=FALSE}
fr_rates %>% 
  filter(strategy == "comprehensive", boundary == "relative") %>% 
  ggplot(aes(a, s, fill = rate)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_tile(colour="white", size=0.25) + 
  scale_x_continuous(expand=c(0,0), breaks = seq(15, 75, 15)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(.1, 1, .1)) +
  scale_fill_viridis() + 
  labs(title = "Comprehensive Integration | Relative Boundary",
       x = "a", 
       y= "s", 
       fill = "% False Responses") + 
  theme_minimal() 
```

#### Switching Probability (s) and Boundary Value (a)

As for both piecewise and comprehensive integration the differences between boundary types are rather minor and of magnitude than of qualitative pattern, the remaining analyses of false response rates are summarized across absolute and relative boundaries.

Below, the $s$ and $a$ parameter are considered as additional sources of variation in the false response pattern above and beyond the interplay of integration strategies and the rarity and attractiveness of outcomes.

```{r message=FALSE}
fr_rates %>% 
  filter(strategy == "piecewise") %>% 
  ggplot(aes(s, rate, color = a)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_jitter(size = 2) + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  scale_color_viridis() + 
  labs(title = "Piecewise Integration",
       x = "s", 
       y= "% False Responses", 
       color = "a") + 
  theme_minimal() 
```

```{r message=FALSE}
fr_rates %>% 
  filter(strategy == "comprehensive") %>% 
  ggplot(aes(s, rate, color = a)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_jitter(size = 2) + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  scale_color_viridis() + 
  labs(title = "Comprehensive Integration",
       x = "s", 
       y= "% False Responses", 
       color = "a") + 
  theme_minimal() 
```

For piecewise integration, switching probability is naturally related to the size of the samples on which the round-wise comparisons of prospects are based on, with low values of $s$ indicating large samples and vice versa.
Accordingly, switching probability is positively related to false response rates.
I.e., the larger the switching probability, the smaller the round-wise sample size and the probability of experiencing a rare event within a given round.
Because round-wise comparisons are independent of each other and binomial distributions within a given round are skewed for small samples and outcome probabilities [@kolmogorovFoundationsTheoryProbability1950], increasing boundary values do not reverse but rather amplify this relation.

For comprehensive integration, switching probability is negatively related to false response rates, i.e., an increase in $s$ is associated with decreasing false response rates.
This relation, however, may be the result of an artificial interaction between the $s$ and $a$ parameter.
Precisely, in the current algorithmic implementation of sampling with a comprehensive integration mechanism, decreasing switching probabilities cause comparisons of prospects based on increasingly unequal sample sizes immediately after switching prospects.
Consequentially, reaching (low) boundaries is rather a function of switching probability and associated sample sizes than of actual evidence for a given prospect over the other.

### Cumulative Prospect Theory

In the following, we examine the possible relations between the parameters of the *choice-generating* sampling models and the *choice-describing* cumulative prospect theory.

For each distinct strategy-parameter combination, we ran 20 chains of 40,000 iterations each, after a warm-up period of 1000 samples.
To reduce potential autocorrelation during the sampling process, we only kept every 20th sample (thinning).

```{r}
# read CPT data
cols <- list(.default = col_double(),
             strategy = col_factor(),
             boundary = col_factor(),
             parameter = col_factor())
estimates <- read_csv("data/estimates/estimates_cpt_pooled_goldstein-einhorn-87.csv", col_types = cols)
```

#### Convergence

```{r}
gel_92 <- max(estimates$Rhat) # get largest scale reduction factor (Gelman & Rubin, 1992) 
```

The potential scale reduction factor $\hat{R}$ was $n \leq$ `r round(gel_92, 3)` for all estimates, indicating good convergence.

#### Piecewise Integration

```{r}
# generate subset of all strategy-parameter combinations (rows) and their parameters (columns)
curves_cpt <- estimates %>% 
  select(strategy, s, boundary, a, parameter, mean) %>% 
  pivot_wider(names_from = parameter, values_from = mean)
```

##### Weighting function w(p)

We start by plotting the weighting curves for all parameter combinations under piecewise integration.

@@

cpt_curves_piecewise <- curves_cpt %>% 
  filter(strategy == "piecewise") %>% 
  expand_grid(p = seq(0, 1, .1)) %>% # add vector of objective probabilities
  mutate(w = round(exp(-delta*(-log(p))^gamma), 2)) # compute decision weights (cf. Prelec, 1998)

# all strategy-parameter combinations

cpt_curves_piecewise %>% 
  ggplot(aes(p, w)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Piecewise Integration: Weighting functions",
       x = "p", 
       y= "w(p)") + 
  theme_minimal() 
```

```{r}
cpt_curves_piecewise %>% 
  ggplot(aes(p, w)) + 
  geom_path() +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  facet_wrap(~a) + 
  labs(title = "Piecewise Integration: Weighting functions",
       x = "p",
       y= "w(p)",
       color = "Switching Probability") + 
  scale_color_viridis() +
  theme_minimal() 
```

```{r}
cpt_curves_piecewise %>% 
  ggplot(aes(p, w, color = s)) + 
  geom_path() +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Piecewise Integration: Weighting functions",
       x = "p", 
       y= "w(p)", 
       color = "Switching Probability") + 
  scale_color_viridis() +
  theme_minimal() 
```

```{r}
cpt_curves_piecewise %>% 
  ggplot(aes(p, w, color = s)) + 
  geom_path() +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  facet_wrap(~a) + 
  labs(title = "Piecewise Integration: Weighting functions",
       x = "p",
       y= "w(p)",
       color = "Switching Probability") + 
  scale_color_viridis() +
  theme_minimal() 
```

##### Value function v(x)

@@

cpt_curves_piecewise <- curves_cpt %>% 
  filter(strategy == "piecewise") %>% 
  expand_grid(x = seq(0, 20, 2)) %>% # add vector of objective outcomes
  mutate(v = round(x^alpha, 2)) # compute decision weights (cf. Prelec, 1998)

# all strategy-parameter combinations

cpt_curves_piecewise %>% 
  ggplot(aes(x, v)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Piecewise Integration: Value functions",
       x = "p", 
       y= "w(p)") + 
  theme_minimal() 
```

```{r}
cpt_curves_piecewise %>% 
  ggplot(aes(x, v, color = s)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Piecewise Integration: Value functions",
       x = "p", 
       y= "w(p)") + 
  scale_color_viridis() + 
  theme_minimal() 
```

```{r}
cpt_curves_piecewise %>% 
  ggplot(aes(x, v, color = s)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  facet_wrap(~a) + 
  labs(title = "Piecewise Integration: Value functions",
       x = "p", 
       y= "w(p)") + 
  scale_color_viridis() + 
  theme_minimal() 
```

#### Comprehensive Integration

##### Weighting function w(p)

We start by plotting the weighting curves for all parameter combinations under piecewise integration.

cpt_curves_comprehensive <- curves_cpt %>% 
  filter(strategy == "comprehensive") %>% 
  expand_grid(p = seq(0, 1, .1)) %>% # add vector of objective probabilities
  mutate(w = round(exp(-delta*(-log(p))^gamma), 2)) # compute decision weights (cf. Prelec, 1998)

# all strategy-parameter combinations

cpt_curves_comprehensive %>% 
  ggplot(aes(p, w)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Comprehensive Integration: Weighting functions",
       x = "p", 
       y= "w(p)") + 
  theme_minimal() 
```

```{r}
cpt_curves_comprehensive %>% 
  ggplot(aes(p, w)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Comprehensive Integration: Weighting functions",
       x = "p", 
       y= "w(p)") + 
  facet_wrap(~a) + 
  theme_minimal() 
```

```{r}
cpt_curves_comprehensive %>% 
  ggplot(aes(p, w, color = s)) + 
  geom_path() +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Comprehensive Integration: Weighting functions",
       x = "p", 
       y= "w(p)", 
       color = "Switching Probability") + 
  scale_color_viridis() +
  theme_minimal() 
```

```{r}
cpt_curves_comprehensive %>% 
  ggplot(aes(p, w, color = s)) + 
  geom_path() +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  facet_wrap(~a) + 
  labs(title = "Comprehensive Integration: Weighting functions",
       x = "p",
       y= "w(p)",
       color = "Switching Probability") + 
  scale_color_viridis() +
  theme_minimal() 
```

```{r}
cpt_curves_comprehensive %>% 
  filter(s >= .7) %>% 
  ggplot(aes(p, w, color = s)) + 
  geom_path() +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  facet_wrap(~a) + 
  labs(title = "Comprehensive Integration: Weighting functions",
       x = "p",
       y= "w(p)",
       color = "Switching Probability") + 
  scale_color_viridis() +
  theme_minimal() 
```

##### Value function v(x)

cpt_curves_comprehensive <- curves_cpt %>% 
  filter(strategy == "comprehensive") %>% 
  expand_grid(x = seq(0, 20, 2)) %>% # add vector of objective outcomes
  mutate(v = round(x^alpha, 2)) # compute decision weights (cf. Prelec, 1998)

# all strategy-parameter combinations

cpt_curves_comprehensive %>% 
  ggplot(aes(x, v)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Comprehensive Integration: Value functions",
       x = "p", 
       y= "w(p)") + 
  theme_minimal() 
```

```{r}
cpt_curves_comprehensive %>% 
  ggplot(aes(x, v)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  facet_wrap(~a) + 
  labs(title = "Comprehensive Integration: Value functions",
       x = "p", 
       y= "w(p)") + 
  theme_minimal() 
```

```{r}
cpt_curves_comprehensive %>% 
  ggplot(aes(x, v, color = s)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  labs(title = "Comprehensive Integration: Value functions",
       x = "p", 
       y= "w(p)") + 
  scale_color_viridis() + 
  theme_minimal() 
```

```{r}
cpt_curves_comprehensive %>% 
  ggplot(aes(x, v, color = s)) + 
  geom_path(size = .5) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  facet_wrap(~a) + 
  labs(title = "Comprehensive Integration: Value functions",
       x = "p", 
       y= "w(p)") + 
  scale_color_viridis() + 
  theme_minimal() 
```

@@

@@

## Appendix 2: Prospect Test Set

@@

>>>>

---
title: 'Sampling Strategies in Decisions from Experience'
author: "Linus Hof, Thorsten Pachur, Veronika Zilker"
bibliography: references_sampling-in-dfe.bib
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
    number_sections: no
  pdf_document:
    toc: yes
csl: apa.csl
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

```{r}
# load packages
pacman::p_load(tidyverse, 
               knitr, 
               scico # for scientific color palettes
               )
```

@@

[...]

@@

[...]

@@

## Generating Model

[...]

The *switching probability* $s$ is the probability with which agents draw a single sample from the prospect they did not get their most recent single sample from.
$s$ is varied between .1 to 1 in increments of .1.
The two boundary parameters resemble the concept of a decision threshold, i.e., if a prospect reaches a boundary, it is chosen by the synthetic agent.
The *boundary type* is either the minimum number of comparisons any prospect must win (absolute boundary) or the minimum difference between the number of won comparisons (relative boundary).
The **boundary value** $a$ is varied between 1 to 5 in increments of 1.

@@

For each parameter combination of the generating model, 100 synthetic agents are presented with 60 choices problems.
In sum, 100 (parameter combinations), 100 (agents) x 60 (choices) = 600,000 choices are simulated.
We test a set of 2-prospect choice problems, where one of the prospects contains a safe outcome, i.e., $p(\omega) = 1$ and the other two outcomes where all $p(\omega) \neq 1$.
Both outcomes and probabilities are drawn from uniform distributions, ranging from 0 to 20 for outcomes and from .01 to .99 for probabilities of the smaller outcome of the risky prospect.
To omit dominant prospects within a choice problem, outcomes of the safe prospect always fall between both outcomes of the risky prospect.
Table A1 contains the test set of 60 choice problems [Appendix 2][Appendix 2: Choice Problems], which were sampled from an initial set of 10,000.
Sampling of gambles was stratified, randomly drawing an equal number of 20 gambles with no, an attractive, and an unattractive rare outcome.
Risky outcomes are considered *"rare"* if their probability is $p < .2$ and *"attractive"* (*"unattractive"*) if they are higher (lower) than the safe outcome.

## Describing Model: Cumulative Prospect Theory

[...]

For each distinct strategy-parameter combination, we ran 20 chains of 20,000 iterations each, after a warm-up period of 1000 samples.
To reduce potential autocorrelation during the sampling process, we only kept every 20th sample (thinning). 
Bar plots displaying the potential scale reduction factors $\hat{R}$ [@gelmanInferenceIterativeSimulation1992] of the group and agent level CPT parameters can be found in [Appendix 3][Appendix 3: Convergence Diagnostics for CPT Parameters].

@@

## Plausibility Check: Relationship between Switching Probability and Trial Sample Sizes

```{r message=FALSE}

# read choice data

cols <- list(.default = col_double(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices <- read_csv("data/choices/choices.csv", col_types = cols)

# get median trial length for each parameter combination

trial_length <- choices %>% 
  group_by(s, boundary, a) %>% 
  summarise(med = median(n_sample))
```

The median length of trials, i.e., the number of single samples drawn in trial, generated by different parameter combinations ranged from `r min(trial_length$med)` to `r max(trial_length$med)`.
As expected, the scatter plot below shows an inverse relationship between switching probability and trial length.
I.e., the lower the switching probability, the larger become the sample sequences on which each comparison between prospects is based.  
This effect is particularly pronounced for low probabilities such that the increase in trial length accelerates as switching probability decreases.

```{r message=FALSE}

# get median trial length for each switching probability

trial_length_s <- choices %>% 
  group_by(s) %>%
  summarise(med = median(n_sample))

# plot

trial_length %>%
  ggplot(aes(x = s, y = med)) +
  geom_jitter(color = "#CECECE", size = 3) +
  geom_point(data = trial_length_s, aes(color = s), size = 3) +
  geom_path(data = trial_length_s, aes(color = s)) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 120, 10)) +
  labs(x ="Switching Probability",
       y = "Median Trial Length", 
       color="Switching Probability") + 
  theme_minimal()
```

## Probability Weighting Function

The following analyses of choice behavior uses the modeling framework of CPT. 
A non parametric approach, analyzing false response rates, can be found in [Appendix 4][Appendix 4: False Response Rates].

@@

# call script for preparing CPT data

source("R/prepare_estimates.R") 
gel_92 <- cpt %>% select(s, boundary, a, parameter, level, Rhat) # store scale reduction factor

# Tidy: CPT parameters in columns

cpt <- cpt %>% 
  select(s, boundary, a, parameter, level, mean) %>% 
  pivot_wider(names_from = parameter, values_from = mean)

# agent level estimates

cpt_agent <- cpt %>% filter(level != "mu") # agent level estimates

## compute mean across all agent level estimates

cpt_agent_mean <- cpt_agent %>% 
  group_by(s, boundary, a) %>% 
  summarise(alpha = round(mean(alpha), 2), 
            gamma = round(mean(gamma), 2), 
            delta = round(mean(delta), 2), 
            rho = round(mean(rho), 2))

# group level estimates

cpt_mu <- cpt %>% filter(level == "mu") # group level posterior means
```

@@

# gamma

cpt_agent %>% 
  ggplot(aes(x = s, y = gamma)) +
  geom_jitter(color = "#CECECE" ) + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 3) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Curvature ", gamma)),
       x ="Switching Probability",
       y = expression(gamma), 
       color="Switching Probability") + 
  theme_minimal()

# delta

cpt_agent %>% 
  ggplot(aes(x = s, y = delta)) +
  geom_jitter(color = "#CECECE") + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 3) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) + 
  scale_y_continuous(breaks = seq(0, 10, .5)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Elevation ", delta)),
       x ="Switching Probability",
       y = expression(delta), 
       color="Switching Probability") + 
  theme_minimal()

# weighting function

## compute decision weights

cpt_agent_w <- cpt_agent %>% 
  select(-c(alpha, rho)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>% # 
  mutate(w = round(exp(-delta*(-log(ep))^gamma), 2))

cpt_agent_mean_w <- cpt_agent %>% 
  group_by(s) %>% 
  summarise(gamma = round(mean(gamma), 2), 
            delta = round(mean(delta), 2)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>%
  mutate(w = round(exp(-delta*(-log(ep))^gamma), 2))

## plot curves

cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w)) +
  geom_path(color = "#CECECE") + 
  geom_path(data = cpt_agent_mean_w, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()

cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability ep",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()

## for separate a

cpt_agent_mean_w <- cpt_agent %>% 
  group_by(s, a) %>% 
  summarise(gamma = round(mean(gamma), 2), 
            delta = round(mean(delta), 2)) %>% 
  expand_grid(ep = seq(0, 1, .1)) %>%
  mutate(w = round(exp(-delta*(-log(ep))^gamma), 2))

cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w)) +
  geom_path(color = "#CECECE") + 
  geom_path(data = cpt_agent_mean_w, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  facet_wrap(~a, nrow = 2) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()

cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w, color = s)) +
  geom_path() + 
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  facet_wrap(~a, nrow = 2) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability ep",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()
```

## Value Function

[...]

@@

[...]

@@

## Appendix 2: Choice Problems

@@

## Appendix 3: Convergence Diagnostics for CPT Parameters

```{r}
# group level

gel_92 %>% 
  filter(level == "mu") %>% 
  ggplot(aes(x = boundary, y = Rhat, fill = s)) +
  geom_bar(position = "dodge_2", stat = "identity") +
  scale_fill_scico(palette = "buda") + 
  scale_x_discrete(name = element_blank()) +
  scale_y_continuous(breaks = seq(.1, 1.5, .2)) +
  geom_hline(yintercept = 1, color = "black") +
  geom_hline(yintercept = 1.1, color = "red", linetype = "dashed") +
  facet_grid(parameter~a, switch = "y") +
  labs(y = "Scale Reduction Factor", 
       fill="Switching Probability") + 
  theme_minimal()

# agent level

parameters <- c("alpha", "gamma", "delta", "rho")
for(i in seq_along(parameters)) {
